Toward single particle reconstruction without particle picking

### 2018-05-06 TEX/draft3.tex
Draft 3

\title{Estimation below the detection limit}

@@

In this paper, we consider the problem of estimating a set of signals $x_1,\ldots,x_K$ from their multiple occurrences in unknown  locations in a data sequence $y$\TODO{This is the most important sentence of the paper, we need to polish it up}. The data may also contain background information---independent of the signals---which we model as noise.
A precise mathematical formulation of the model and the estimation problem is provided in Section~\ref{sec:model}.
For one-dimensional signals, the data sequence can be thought of as a 
long time series and the $K$ signals as repeated short events.  
This model appears in many applications, including spike sorting~\cite{lewicki1998review}, passive radar~\cite{gogineni2017passive} and system identification~\cite{ljung1998system}.
In the last part of this paper, we also propose to interpreted this estimation problem as a toy model for the task of single particle reconstruction using cryo-electron microscopy (cryo--EM).

Figure~\ref{fig:XC_example} shows a simple example of data sequences with  one signal ($K=1$).
The upper row presents the data without noise, in the left panel, and its cross-correlation with the signal itself in the right panel. The cross-correlation exhibits two peaks, corresponding to the two signal's locations in the data. 
%Even if we add a small amount of noise and considering more than one signal ($K>1$), estimating the signals remains a painless task. 
%In this scenario, standard detection and clustering algorithms can locate and cluster the copies of the signals that can be then averaged within each class.  
In  the middle row, the same data is shown, now contaminated with i.i.d.\ Gaussian noise with zero mean and standard deviation of $\sigma=0.5$. 
While it is harder to identify the signal occurrences in the left panel, the cross-correlation still produces fairly adequate estimates. In practice one usually does not possess a precise template of the signal to be recovered, however, clever methods based on template matching, such as those used in structural biology~\cite{heimowitz2018apple} and radar~\cite{gogineni2017passive}, may work.
The bottom panels show the same data swamped in  
i.i.d.\ Gaussian noise with standard deviation  $\sigma=3$. In this  low signal--to--noise (\SNR) regime, 
detection of individual signal occurrences is impossible, 
even if the true signal is known. 
This phenomenon---explained in more detail in Section~\ref{sec:model}--raises the following fundamental question: 
%\medskip 
%\\
\begin{center}
\emph{Can we estimate the signals accurately without detecting their individual occurrences? }  %\\
\end{center}

In this work, we provide a positive answer to this question. For a single signal ($K=1$), we prove that in the asymptotic regime in which the signal appears
infinitely many times, and under a spacing condition, one can estimate the signal \emph{to any desired accuracy in any $\SNR$ level}. We show empirically that the same holds true for multiple signals if $K$ is not too large. \TODO{Do we want a sentence like: Based on  recent results on the multireference alignment problem, we conjecture that it remains true as long as $K \lesssim L/6$. ? I am not sure it is necessary.}

Our framework is based on autocorrelation analysis.
In a nutshell, the method consists of two stages. First, we estimate a mixture (i.e., linear combination) of the low--order autocorrelation functions of the signals from the data. These quantities can be estimated, to any desired accuracy, if individual occurrences are separated by at least $L-1$ entries and  each signal appears sufficiently many times in the data. There is no need to detect individual occurrences. 
Furthermore, we do not assume the knowledge of the number of signal occurrences $M$ or the noise level.
In the second stage, the signals are estimated from the mixed autocorrelations using a nonconvex least-squares (LS).
Section~\ref{sec:autocorrelation} elaborates on the technique and proves some of its properties, while Section~\ref{sec:numerics} shows numerical demonstrations and discusses computational aspects.

Interestingly, expectation-maximization (EM)---a popular algorithm for similar estimation problems, such as Gaussian mixture models and multireference alignment---is intractable for this problem. This is true even if  $K=1$ and the number of signal occurrences $M$ is known.
In particular, in each iteration, EM assigns a probability to any feasible combination of positioning the current signal estimate in $M$ locations on the grid $\{1,\ldots,N\}$.
In total, even when excluding forbidden combinations due to the spacing constraint, there are $O(N^M)$ such combinations. \TODO{cumbersome}

\begin{figure}
		\centering
		\includegraphics[scale=0.5]{XC_example}
		\caption{ The upper left panel shows a data sequence (measurement) of length $N=80$ with one rectangular signal of length $L=11$ that appears twice. The upper right panel shows the cross-correlation of the data with the true rectangular signal.  
		The second and third rows present the data sequence with additive Gaussian noise with mean zero and standard deviation $\sigma=0.5$ and $\sigma=3$, respectively, and its cross-correlation with the signal. Note to the different scales in different panels. Importantly, in the low $\SNR$ regime the cross-correlation does not provide information on the locations of the signal in the data.}
    	\label{fig:XC_example}
\end{figure}

%
%This work is primely motivated by cryo-electron microscopy (cryo--EM), which is an innovative  single particle reconstruction technology. The acquired data in a cryo-EM experiment is contaminated with high noise levels. Therefore, any molecule reconstruction algorithm must take  the challenging $\SNR$ level into account.  In the last part of this manuscript, we draw connections with the estimation problem under consideration and the cryo--EM problem.

\section{Model and related literature}  \label{sec:model}

Let $x_1,\ldots,x_K\in\RL$ be the sought signals and let $y\in\RN$ be the data. 
The forward model can be posed as a mixture of \emph{blind deconvolution} problems between binary signals and the target signals $x_i$:
\begin{equation} \label{eq:model}
y = \sum_{i=1}^K x_i\ast s_i + \varepsilon,\quad \varepsilon\sim\mathcal{N}(0,\sigma^2 I).
\end{equation}
%We model the background information as i.i.d.\ Gaussian noise with zero mean and $\sigma^2$ variance. 
The nonzero values of each  $s_i\in\{0,1\}^N$ determine the position of the occurrences of the corresponding $x_i$. 
We denote the set of these nonzero values by 
 $\mathcal{S}_i$ and its cardinality by $\vert \mathcal{S}_i\vert = M_i$. 
By assuming that all $\mathcal{S}_i$'s are disjoint,  we let $s = \sum_{i=1}^Ks_i$, $\mathcal{S} = \bigcup_{i=1}^{K} \mathcal{S}_i$ and  $\vert \mathcal{S}\vert :=M =  \sum_{i=1}^{K}M_i$.  Neither the $M_i$'s nor $M$ are assumed to be known.

Blind deconvolution is a longstanding problem, arising in a variety of engineering and scientific applications, such as astronomy, communication, image deblurring, system identification and optics; see~\cite{jefferies1993restoration,shalvi1990new,ayers1988iterative,abed1997blind}, just to name a few. 
To make the problem well-posed, we must  assume some prior knowledge or  structure. 
In our case, the prior information is that $s$ is a binary signal that satisfies the separation constraint~\eqref{eq:spacing}. 
Other settings of blind deconvolution problems have been analyzed recently, see for instance~\cite{ahmed2014blind,li2016identifiability,li2016rapid,ling2015self,ling2017blind,chi2016guaranteed}
where the focus is on high $\SNR$ regimes.

If $x$ is known and $K=1$, then a sparse signal can be estimated by linear programming  in the high $\SNR$ regime, e.g.,~\cite{azais2015spike,denoyelle2017support,bendory2016robust,bendory2017robust,bernstein2017deconvolution}. However, in the low $\SNR$ regime, estimating the binary sparse signal $s$ is impossible. To see that, suppose that an oracle provides us $M$ windows of length $W>L$, each contains one copy of $x$. That is to say, we get a series of windows of length $W$, each one contains a signal at an unknown location. 
Estimating the position of the known signal within each  window is an easier problem than detecting the support of $s$. 
Nevertheless, even this problem is impossible in the low $\SNR$ regime~\cite{aguerrebere2016fundamental}. Therefore, we conclude that detecting the nonzero values of $s$ is impossible in low $\SNR$. As aforementioned, this work focuses on this regime and examines under what conditions we can estimate the signals, despite the impossibility of detecting their individual occurrences.

For $K=1$, our problem can be also interpreted as a special case of the system identification problem. Similarly to~\eqref{eq:model}, the system identification forward model takes the form
\begin{eqnarray}
y = x\ast w + \varepsilon,  
\end{eqnarray} 
where $x$ is the unknown signal (``system''), $w$ is an unknown, random, input sequence and $\varepsilon$ is an additive noise.   
%The problem has also been studied in the case of a known input $w$~\cite{pillonetto2010new,dinuzzo2015kernels,bottegal2016robust}. 
The goal of this problem is to estimate $x$, usually referred to as ``identifying the system.'' The question of identifiability of $x$ under
this observation model is addressed for certain Gaussian and non-Gaussian $w$ in~\cite{benveniste1980robust,kormylo1983identifiability}.
In the special case where $w\in\{0,1\}^N$, satisfying the spacing requirement~\eqref{eq:spacing}, we obtain our
model in the  case of a single signal ($K = 1$).

\section{Autocorrelation analysis}   \label{sec:autocorrelation}

Our method for estimating the signals is composed of two stages. 
First, we use the autocorrelation functions of the data to estimate a mixture (i.e., linear combination) of the $K$ signal's autocorrelations. The mixed autocorrelation can be estimated to any accuracy, in any $\SNR$ level, if $M$ is large enough and the spacing condition~\eqref{eq:spacing} is met. Then, we  use a nonconvex LS  to estimate the signals from their mixed autocorrelations. 
In this section, we elaborate on the autocorrelation functions and their estimations, while the precise recovery procedure, based on nonconvex optimization, will be discussed in detail in the next section.

\subsection{Aperiodic autocorrelation functions} \label{sec:aperiodic_ac}

For the purpose of this paper, we need the first three (aperiodic) autocorrelation functions. The first-order autocorelation is the mean of the signals. For  
$z\in\RL$ and $k\geq 2$, the autocorrelation of order $k$ is defined for any integer shifts $\ell_1, \ldots, \ell_{k-1}$ by
\begin{align}
	a_z^k[\ell_1,\ldots,\ell_{k-1}]  & = \sum_{i=-\infty}^{+\infty} z[i]z[i+\ell_1]\ldots z[i+\ell_{k-1}],
	\label{eq:ac_general}
\end{align}
where indexing of $z$ out of the bounds $0, \ldots, L-1$ is zero-padded, as usual.
Explicitly, the first three autocorrelations are
\begin{align} 
	a_z^1 & = \sum_{i=0}^{L-1} z[i], \nonumber\\
	a_z^2[\ell] & = \sum_{i = \max\{0, -\ell\}}^{L-1 + \min\{0, -\ell\}} z[i]z[i+\ell], \nonumber\\
	a_z^3[\ell_1,\ell_2] & = \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{L-1 + \min\{0, -\ell_1, -\ell_2\}} z[i]z[i+\ell_1]z[i+\ell_2]. \label{eq:ac_special}
\end{align}
Note that the autocorrelation functions are symmetric so that $a_z^2[\ell] = a_z^2[-\ell]$ and $$a_z^3[\ell_1,\ell_2] = a_z^3[\ell_2,\ell_1]=a_z^3[-\ell_1,\ell_2-\ell_1].$$
Additionally, if the moments of the signal depend only on the difference between the indices (Toeplitz structure), then they are equivalent to the autocorrelation functions.

A few remarks are in order. 
First, the second result of Proposition~\ref{prop:uniqueness} shows that there exists a very simple estimator that has finite sensitivity. In the next section, we propose an estimator based on nonconvex LS that shows empirical robustness to additive noise, in accordance with related problems~\cite{bendory2017bispectrum,boumal2017heterogeneous}. 
Second, these results carry through to signals of any dimension.
Third, if the spacing condition~\eqref{eq:spacing} holds, then the length of the signal can be determined from the autocorrelations and 
therefore the assumption that the first and last entries are nonzero is met. In particular, if~\eqref{eq:spacing} holds for some spacing $W\geq L$, then $a_z^2[i]=0$ for all $i>L-1$.
Finally, computing the $d$th autocorrelation amplifies the variance of the noise by a factor $d$ in the low $\SNR$ regime. Therefore, if we can estimate $a_z^3$ up to small perturbation, it implies that we can estimate $a_z^2$ accurately as the proposition assumes.

\TODO{NB: I would remove most if this paragraph; we can have a similar discussion but for our case (after having discussing which moments we keep); make the statement, and finish with a one-sentence reference to the MRA paper. TB: I didn't touch it yet. We may want to put the degrees-of-freedom calculations from Section 4 here}
The invertibility of the autocorrelations for $K>1$ was explored  for the related case of  periodic autocorrelation functions. Recall that the  periodic autocorrelations are defined similarly to~\eqref{eq:ac_general}, with two differences: the sum goes to $L-1$ and all indices should be taken modulo $L$.
If $z[n]=0$ for $n=L/2,\ldots L-1$ (``zero-padded''), then the periodic and aperiodic autocorrelation coincide.
In~\cite{bandeira2017estimation}, it was shown that a mix of $K$ third-order periodic autocorrelations determine  a finite list of $K$ generic signals if $L/6\gtrsim K$. Empirical evidences hints that this finite list includes only group symmetries, at least as long as $K\leq\sqrt{L}$~\cite{boumal2017heterogeneous}.

\subsection{Estimating autocorrelations from the data with a single signal $K=1$} \label{sec:estimating_ac_k1}

We first consider the problem of estimating the autocorrelations of a single signal from the data. 
The main principles carry through for $K>1$ as will be shown in the next section.

In order to estimate the autocorrelations of the signal, we first compute the first $L$ entries of the data's autocorrelations. 
For the purpose of the analysis, we consider  the asymptotic regime where $M,N\to\infty$, while preserving fixed ratio. 
Specifically, we define the ratio of the measurement occupied by  the signal as
\begin{equation}
\gamma = \frac{M L}{N}.
\end{equation}
Under the spacing constraint~\eqref{eq:spacing}, we have $\gamma\leq\frac{L}{2L-1}\approx 1/2$.

The main pillar of this work is the following simple observation.
If the support signal $s$ satisfies the spacing constraint~\eqref{eq:spacing}, then the first $L$ entries of the data autocorrelations converge 
to a scaled, biased, version of the signal's autocorrelation:
\begin{align}
\lim_{N\to\infty} a_y^1 & = \gamma a_{x}^1, \nonumber\\
\lim_{N\to\infty} a_y^2[\ell] & = \gamma a_{x}^2[\ell] +\sigma^2\delta[\ell], \label{eq:data_ac_k1} \\
\lim_{N\to\infty} a_y^3[\ell_1,\ell_2] & = \gamma a_{x}^3[\ell_1,\ell_2] + \sigma^2\gamma a_{x}^1 \big(\delta[\ell_1,0]+\delta[0,\ell_2]+\delta[\ell_1,\ell_2]\big), \nonumber
\end{align}
for $\ell,\ell_1,\ell_2=0,\ldots L-1$, and where $\delta$ denotes the Kronecker delta function. 
These relations are proven in Appendix~\ref{sec:autocorrelation_computation}. The analysis is similar to~\cite{bendory2017bispectrum,boumal2017heterogeneous}, yet a particular caution should be taken with the statistical dependencies of the noise entries. 
The relations~\eqref{eq:data_ac_k1}, together with Proposition~\ref{prop:uniqueness}, imply that given $M$ and $\sigma$, one can estimate the signal for any noise level if $M$ is large enough. Next, we show that  $M$ and $\sigma$ are also uniquely determined from the autocorrelations.

\subsection{Estimating autocorrelations from the data with multiple signals $K>1$} \label{sec:estimating_ac}

The minimal order of data statistics used to get an accurate estimation of a signal is important to understand, in the asymptotic $\SNR$ regime, the sample complexity of the problem.
In methods which are based on detection and averaging, the number of signals occurrences  must scale like $\sigma^2$. Taking the $d$th order autocorrelation function amplifies the variance of the noise by a factor of $d$. Therefore, the required number signal occurrences should scale like $\sigma^{2d}$  to retain some constant estimation error. Accordingly, in our  method, $M$ must scale like $\sigma^6$. In Section~\ref{sec:numerics}, we show empirically the third-order autocorrelation suffices also for a mixture of $K>1$ signals without prior knowledge of the $M_i$'s and $\sigma$.  \TODO{question mark on the entire paragraph}
\TODO{NB: We could move here the paragraphs from 1D XP about which moments we compute, to argue that we can expect to recover at most up to $K = L/2$ signals.} \TODO{TB: I didn't touch for now.}

\section{Numerical experiments}   \label{sec:numerics}

@@

%\subsection{One-dimensional experiments}

\TODO{Revise to explain $K = 1$ experiment first, then explain $K = 3$.}

\begin{figure}[t]
	\centering
	\includegraphics[width=0.7\linewidth]{XP_1D_homogeneous_new_ROI/progressive_n10000000000_107521}
	\caption{\TODO{...} \TODO{This figure is with new ROI method based on loss functions}}
	\label{fig:1Dhomosignals}
\end{figure}

@@

@@

Given the observation $y$, we proceed to compute the moments. The first-order moment is straightforward. For second-order moments, notice from equation~\eqref{eq:data_ac} that $a_y^2[\ell]$ suffers no bias for $\ell$ in $1$ to $L-1$. Thus, we omit $\ell = 0$, which has the practical effect that we need not know $\sigma$ to estimate the moments. Likewise, for third-order moments, $a_y^3[\ell_1, \ell_2]$ for $0 \leq \ell_1, \ell_2 \leq L-1$ such that $\ell_2 \leq \ell_1$ includes all relevant moments for our purpose \TODO{Do we want to include a figure to explain why that is?}, and we further exclude any such that $\ell_1, \ell_2$ or $\ell_1 - \ell_2$ are zero to avoid biased elements---there are $\frac{(L-1)(L-2)}{2}$ remaining moments. As a result, it is unnecessary to estimate $\sigma$. We have \TODO{TB: We may want to put this calculation is the last paragraph of Section 3.1 }
\begin{align*}
	1 + (L-1) + \frac{(L-1)(L-2)}{2} = \frac{1}{2} L (L-1) + 1
\end{align*}
moments in total. In practice, these are computed on disjoint segments of $y$ of length $100\cdot10^6$ and added up, without correction for the junction points. Segments are handled sequentially on a GPU, as GPUs are particularly well suited to execute simple instructions across large vectors of data. If multiple GPUs are available, segments can of course be handled in parallel.

@@

\TODO{TB: 2D experiment?}
%\subsection{Two-dimensional experiments} 
%??

>>>>

\title{Toward single particle reconstruction without particle picking}

@@

Single particle reconstruction (SPR) using cryo--electron microscopy (cryo--EM) is an innovative technology for reconstructing the 3-D structure of macromolecules. In recent years, structures
of many molecules, previously regarded as insurmountable, are now being reconstructed to near-atomic resolution; see for instance~\cite{kuhlbrandt2014resolution,bartesaghi20152}. This technological advancement was recognized by the 2017 Nobel Prize in Chemistry~\cite{nobel}.

In a cryo--EM experiment, multiple biological samples of the (ideally) same molecule are rapidly frozen in a thin layer of vitreous ice. Within the ice, the molecules are randomly oriented and positioned. Then, the microscope produces a 2D tomographic projection image, called a \emph{micrograph}, of the multiple samples embedded in the ice. The micrograph is dominated by noise due to the small electron doses that
can be applied to the specimen without causing radiation damage.
The cryo--EM problem is to estimate the structure of the molecule from the micrograph (or, typically, several micrographs).

\paragraph{Particle picking.}
All contemporary methods in the field split the reconstruction procedure to several subroutines. 
The first step in the pipeline is the so-called \emph{particle picking}, in which one aims to detect the 2D tomographic projections of the samples from the noisy micrograph. The output of ideal particle picker is a series of 2D images, each contains one centered tomographic projection associated with an unknown 3D orientation. This series of images 
is then used to estimate the structure.

Many automatic and semi-automatic methods for particle picking have been proposed, based on edge detection, template matching and deep learning; see for instance~\cite{harauz1989automatic,ogura2004automatic,zhu2016deep,frank1983automatic,scheres2015semi,heimowitz2018apple}. 
However, most of these procedures are prone to \emph{model bias}. For instance, in the popular framework of RELION~\cite{scheres2015semi}, the user manually marks hundreds of spots on the micrograph, believed to contain projections. 
Therefore, the algorithm performance depends on the user prior assumption about the particle's structure; the same holds true for deep learning based approaches which require constructing labeled set of data.
Other methods use disks or difference of Gaussians as templates~\cite{langlois2014automated,voss2009dog}.
Nowadays, it also still popular to pick particles manually. This method, while exploits the researcher's experience, is both tedious and subject to model bias.

The significance of the model bias in cryo--EM was stressed in~\cite{shatsky2009method,henderson2013avoiding} and demonstrated by the ``Einstein from noise'' example.
In this experiment, the image of Einstein is correlated with multiple images of pure noise. The noisy images are then aligned to Einstein's image using cross-correlation and  averaged. 
In~\cite{shatsky2009method}, it was shown that using merely 1000 noise images, the averaged image is remarkably similar to the template image of 
Einstein, rather than to pure noise. 
%Therefore, such a simple algorithm is biased towards its model which is an undesirable property. 
In the context of cryo--EM, this example demonstrates how our prior assumptions about the particle may influence the reconstructed structure.

Besides the model bias, particle pickers have several limitations due to the high noise level, in particular for small particles and in low contrast. 
As a result, the projections in the 2D images are typically not centered, increasing dramatically the number of parameters involved in the reconstruction algorithm. In addition, the information from particles that are too close to each other is usually neglected. Hence, valuable information that can be harnessed is omitted.

\paragraph{Simplified model for SPR} In this paper, we present a preliminary study that aims showing that estimating a particle 
directly from the micrograph might be possible.

In particular, we consider the problem of estimating a set of signals $x_1,\ldots,x_K$ from their multiple occurrences in unknown  locations in a noisy data sequence $y$. Here, $y$ plays the role of the micrograph and the $K$ signals can be thought of as $K$ different viewing directions of the particle that appear multiple times in the micrograph. 
A precise mathematical formulation of the model and the estimation problem is provided in Section~\ref{sec:methods}.
To be clear, we do not consider here many prominent features of SPR experiment and do not aim to reconstruct any 3-D structure. We also mention by passing that this model appears in different scientific fields, such as spike sorting~\cite{lewicki1998review}, passive radar~\cite{gogineni2017passive} and system identification~\cite{ljung1998system}.

Figure~\ref{fig:micro_example} shows a simple example of our problem with  one signal ($K=1$). 
Each panel demonstrates an image of size $250\times 250$, containing 4 occurrences of an Einstein image of size $50\times 50$. 
The left panel shows the data without noise. In the middle panel, an i.i.d.\ Gaussian noise with standard deviation $\sigma=0.5$  was added, however,  Einstein is still easily identified. In this scenario, even for $K>1$, it is likely that standard detection and clustering algorithms can locate and cluster the copies of the signals that can be then averaged within each class.  In the right panel, we show the same data swamped in  
i.i.d.\ Gaussian noise with standard deviation  $\sigma=3$.
As a matter of fact, it can be shown that in  low signal--to--noise (\SNR) regime, 
detection  of individual image occurrences is impossible---even if the true image is known---and therefore particle picking is impossible.  This phenomenon is explained in detail in Section~\ref{sec:related_literature}.

\begin{figure}[h!]
	\centering
	\begin{subfigure}[h]{0.33\textwidth}
		\centering
		\includegraphics[scale=0.5]{micrograph_Einstein_example_clean}
		\caption{$\sigma = 0$}
	\end{subfigure}%
	\begin{subfigure}[h]{0.33\textwidth}
		\centering
		\includegraphics[scale=0.5]{micrograph_Einstein_example_s05}
		\caption{$\sigma = 0.5$}
	\end{subfigure}
	\begin{subfigure}[h]{0.33\textwidth}
		\centering
		\includegraphics[scale=0.5]{micrograph_Einstein_example_s3}
		\caption{$\sigma = 3$}
	\end{subfigure}
	\caption{\label{fig:micro_example} Example of data sequences (micrographs) of size $250\times 250$ with additive i.i.d./ Gaussian noise with variance $\sigma^2$. Each micrograph contains four occurrences of a $50 \times 50$ image of Einstein.}	
\end{figure}

\paragraph{Autocorrelation analysis.}
In order to recover the signal from the data, we use autocorrelation analysis.
In a nutshell, the method consists of two stages. First, we estimate a mixture (i.e., linear combination) of the low--order autocorrelation functions of the signals from the data. These quantities can be estimated, to any desired accuracy, if individual occurrences are separated by the size of the signal and  each signal appears sufficiently many times in the data. There is no need to detect individual occurrences, knowing the number of signal occurrences $M$ or the noise level.
In the second stage, the signals are estimated from the mixed autocorrelations using a nonconvex least-squares (LS) or phase retrieval algorithms; see Section~\ref{sec:methods}.
This method requires only one pass over the data and can recover the signals in any $\SNR$ level, if the signals appear enough times in the data. As a side note, we mention that expectation-maximization---a popular framework in SPR---is intractable for this problem (see Appendix~\ref{sec:theory}  for more details).

\section{Results}

@@

\TODO{Revise to explain $K = 1$ experiment first, then explain $K = 3$.}
\TODO{TB: we should remove the ``long estimation" from the plots}

In this section we show numerical results, while leaving the specific details of the experiments for Section~\ref{sec:methods}. In the first experiment, we estimated Einstein's image from micrographs with $\sigma=3$ . An example for a micrograph appears in the right panel of Figure~\ref{fig:micro_example}. As initial guess, we picked Newton image. If the algorithm was prone to model bias, we would expect to get as an output an image that resembles Newton, similarly to the ``Einstein from noise'' effect. Nonetheless, the more data we collect, the better the reconstruction.\TODO{We have a movie in the supplementary material.}

\begin{figure}[h!]
	\centering
	\begin{subfigure}[h]{0.25\textwidth}
		\centering
		\includegraphics[scale=0.4]{Newton}
		\caption{Newton  (model)}
	\end{subfigure}%
	\begin{subfigure}[h]{0.25\textwidth}
		\centering
		\includegraphics[scale=0.4]{Einstein_100}
		\caption{$P =10^2$}
	\end{subfigure}%
	\begin{subfigure}[h]{0.25\textwidth}
		\centering
		\includegraphics[scale=0.4]{Einstein_1e3}
		\caption{$P =10^3$}
	\end{subfigure}%
	\begin{subfigure}[h]{0.25\textwidth}
		\centering
		\includegraphics[scale=0.4]{Einstein_1e4}
		\caption{$P = 10^4$}
	\end{subfigure}%
	\caption{\label{fig:Einst_example} Recovery of Einstein image. The algorithm is initialized by the image of Newton. We show estimations of Einstein using $P=10^2,10^3$ and $10^4$ number of micrographs, each contains $860$ image occurrences on average. The normalized recovery error is $0.555$, $0.3988$ and $0.2907$, respectively. The normalized error in recovering the power spectrum was $69.1\times10^{-3} $, $21.7\times10^{-3} $ and $6.8\times10^{-3}$, respectively. \TODO{This image will be more impressive in the future}}	
\end{figure}

Here we should describe the 1-D experiments.

\begin{figure}[t]
	\centering
	\includegraphics[scale=0.6]{XP_1D_homogeneous_new_ROI/progressive_n10000000000_107521}
	\caption{\TODO{...} \TODO{This figure is with new ROI method based on loss functions}}
	\label{fig:1Dhomosignals}
\end{figure}

@@

\section{Discussion}

All current SPR algorithms start with a particle picking stage which is prone to model bias. 
In this paper, we propose to examine the possibility to bypass particle picking and estimate the particle directly from the micrograph. To this end, we examine a simplified model that share some similarities with the SPR problem. The proposed reconstruction method is based on autocorrelation analysis, 
somewhat similar to a technique proposed by Kam which can be used for ab inito modeling~\cite{kam1980reconstruction,levin20173d,singer2018mathematics}. 
That being said, the SPR model is far more complicated thans the model presented here. In a future research, we hope to bridge this gap.

Our results rely on two core assumptions that are not necessarily met by in an SPR experiment.  
First, we modeled the background information as i.i.d.\ additive noise. In practice, the background information may be structured or depend on the signal. However, methods based on autocorrelation estimation are quite flexible and in many cases can tolerate mismatch of the noise statistical model.

In addition, we assumed that the signal occurrences are all separated, see~\eqref{eq:spacing}. 
This separation can be induced by careful experiment design \TODO{???}. 
If the signals are not separated, one can introduce a new variable that represents the distribution of the spacing between signal occurrences and then compute explicitly the relation between the autocorrelation of the data and the signals. It is yet to be studied under what conditions on this distribution one can estimate the signals.

%The first entry $p[1]$ will represent the probability that two consecutive signals are separated by only one entry, $p[2]$ the probability for spacing of two entries and so on. Using this auxiliary variable $p$, one can write explicitly the relation between the autocorrelation functions of the data and those of the signal in a similar way to~\eqref{eq:data_ac}. 
%An interesting question is  under what conditions on $p$ and the signals, one can estimate the signals from the data.

\section{Methods} \label{sec:methods}

\paragraph{Mathematical model}

Let $x_1,\ldots,x_K\in\RL$ be the sought signals and let $y\in\RN$ be the data. 
The forward model can be posed as a mixture of \emph{blind deconvolution} problems between binary signals and the target signals $x_i$:
\begin{equation} \label{eq:model}
y = \sum_{i=1}^K x_i\ast s_i + \varepsilon,\quad \varepsilon\sim\mathcal{N}(0,\sigma^2 I).
\end{equation}
%We model the background information as i.i.d.\ Gaussian noise with zero mean and $\sigma^2$ variance. 
The nonzero values of each  $s_i\in\{0,1\}^N$ determine the position of the occurrences of the corresponding $x_i$. 
We denote the set of these nonzero values by 
 $\mathcal{S}_i$ and its cardinality by $\vert \mathcal{S}_i\vert = M_i$. 
By assuming that all $\mathcal{S}_i$'s are disjoint,  we let $s = \sum_{i=1}^Ks_i$, $\mathcal{S} = \bigcup_{i=1}^{K} \mathcal{S}_i$ and  $\vert \mathcal{S}\vert :=M =  \sum_{i=1}^{K}M_i$.  Neither the $M_i$'s nor $M$ are assumed to be known. Literature survey on blind deconvolution and related problems is given in Appendix~\ref{sec:related_literature}.

%
%\section{Autocorrelation analysis}   \label{sec:autocorrelation}
%
%Our method for estimating the signals is composed of two stages. 
%First, we use the autocorrelation functions of the data to estimate a mixture (i.e., linear combination) of the $K$ signal's autocorrelations. The mixed autocorrelation can be estimated to any accuracy, in any $\SNR$ level, if $M$ is large enough and the spacing condition~\eqref{eq:spacing} is met. Then, we  use a nonconvex LS  to estimate the signals from their mixed autocorrelations. 
%In this section, we elaborate on the autocorrelation functions and their estimations, while the precise recovery procedure, based on nonconvex optimization, will be discussed in detail in the next section.

\paragraph{Aperiodic autocorrelation functions}
%\subsection{Aperiodic autocorrelation functions} \label{sec:aperiodic_ac}

For the purpose of this paper, we need the first three (aperiodic) autocorrelation functions. The first-order autocorelation is the mean of the signals. For  
$z\in\RL$ and $k\geq 2$, the autocorrelation of order $k$ is defined for any integer shifts $\ell_1, \ldots, \ell_{k-1}$ by
\begin{align}
	a_z^k[\ell_1,\ldots,\ell_{k-1}]  & = \sum_{i=-\infty}^{+\infty} z[i]z[i+\ell_1]\ldots z[i+\ell_{k-1}],
	\label{eq:ac_general}
\end{align}
where indexing of $z$ out of the bounds $0, \ldots, L-1$ is zero-padded, as usual.
Explicitly, the first three autocorrelations are
\begin{align} 
	a_z^1 & = \sum_{i=0}^{L-1} z[i], \nonumber\\
	a_z^2[\ell] & = \sum_{i = \max\{0, -\ell\}}^{L-1 + \min\{0, -\ell\}} z[i]z[i+\ell], \nonumber\\
	a_z^3[\ell_1,\ell_2] & = \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{L-1 + \min\{0, -\ell_1, -\ell_2\}} z[i]z[i+\ell_1]z[i+\ell_2]. \label{eq:ac_special}
\end{align}
Note that the autocorrelation functions are symmetric so that $a_z^2[\ell] = a_z^2[-\ell]$ and $$a_z^3[\ell_1,\ell_2] = a_z^3[\ell_2,\ell_1]=a_z^3[-\ell_1,\ell_2-\ell_1].$$
%Additionally, if the moments of the signal depend only on the difference between the indices (Toeplitz structure), then they are equivalent to the autocorrelation functions.

\paragraph{Estimating the autocorrelation function of a single signal}

We first consider the problem of estimating the autocorrelations of a single signal from the data. 
The main principles carry through for $K>1$ as will be shown next.

%In order to estimate the autocorrelations of the signal, we first compute the first $L$ entries of the data's autocorrelations. 
For the purpose of the analysis, we consider  the asymptotic regime where $M,N\to\infty$, while preserving fixed ratio. 
Specifically, we define the ratio of the measurement occupied by  the signal as
\begin{equation}
\gamma = \frac{M L}{N}.
\end{equation}
Under the spacing constraint~\eqref{eq:spacing}, we have $\gamma\leq\frac{L}{2L-1}\approx 1/2$.

The main pillar of this work is the following simple observation.
If the support signal $s$ satisfies the spacing constraint~\eqref{eq:spacing}, then the first $L$ entries of the data autocorrelations converge 
to a scaled, biased, version of the signal's autocorrelation:
\begin{align}
\lim_{N\to\infty} a_y^1 & = \gamma a_{x}^1, \nonumber\\
\lim_{N\to\infty} a_y^2[\ell] & = \gamma a_{x}^2[\ell] +\sigma^2\delta[\ell], \label{eq:data_ac_k1} \\
\lim_{N\to\infty} a_y^3[\ell_1,\ell_2] & = \gamma a_{x}^3[\ell_1,\ell_2] + \sigma^2\gamma a_{x}^1 \big(\delta[\ell_1,0]+\delta[0,\ell_2]+\delta[\ell_1,\ell_2]\big), \nonumber
\end{align}
for $\ell,\ell_1,\ell_2=0,\ldots L-1$, and where $\delta$ denotes the Kronecker delta function. 
%These relations are proven in Appendix~\ref{sec:autocorrelation_computation}.

\paragraph{Estimating the autocorrelation function of a multiple signals}

\paragraph{Numerical experiments details}

\TODO{We may want to split it into two sections. Here we could give a short description with the LS estimator and the RRR. In the supplementary material we can provide all other details. }

@@

\TODO{TB: I would place the equation counting argument in the theory section in the paragraph of open questions (I marked th place)}
Given the observation $y$, we proceed to compute the moments. The first-order moment is straightforward. For second-order moments, notice from equation~\eqref{eq:data_ac} that $a_y^2[\ell]$ suffers no bias for $\ell$ in $1$ to $L-1$. Thus, we omit $\ell = 0$, which has the practical effect that we need not know $\sigma$ to estimate the moments. Likewise, for third-order moments, $a_y^3[\ell_1, \ell_2]$ for $0 \leq \ell_1, \ell_2 \leq L-1$ such that $\ell_2 \leq \ell_1$ includes all relevant moments for our purpose \TODO{Do we want to include a figure to explain why that is?}, and we further exclude any such that $\ell_1, \ell_2$ or $\ell_1 - \ell_2$ are zero to avoid biased elements---there are $\frac{(L-1)(L-2)}{2}$ remaining moments. As a result, it is unnecessary to estimate $\sigma$. We have \TODO{TB: We may want to put this calculation is the last paragraph of Section 3.1 }
\begin{align*}
1 + (L-1) + \frac{(L-1)(L-2)}{2} = \frac{1}{2} L (L-1) + 1
\end{align*}
moments in total. In practice, these are computed on disjoint segments of $y$ of length $100\cdot10^6$ and added up, without correction for the junction points. Segments are handled sequentially on a GPU, as GPUs are particularly well suited to execute simple instructions across large vectors of data. If multiple GPUs are available, segments can of course be handled in parallel.

@@
